Coalition Formation Algorithm of Prosumers in a Smart Grid Environment

### 2014-09-18 Article.tex
update captions and few modif in V

\begin{equation}
 \mathcal{U}_{\phi,\ P_{\phi}^{MIN}}(S) = \mathbf{1}_{\textit{S\ valid}} \dfrac{P_{\phi}(S)}{|S|} 
\label{utility}
\end{equation}

@@

\begin{figure}
 \centering
  \subcaptionbox{$ (\phi, P^{MIN}) $ subspace\label{fig2:a}}{\includegraphics[width=1.7in]{./figure1/util}}
  \subcaptionbox{$ (N_{COAL}, P^{MIN}) $ subspace\label{fig2:b}}{\includegraphics[width=1.70in]{./figure6/util2}}
\caption{Utility function as function of the parameters space $\{\phi \in [0,1], N_{COAL}\in \mathbb{Z}^+, P^{MIN} \in \mathbb{R}^+\}$: In $ \{\phi, P^{MIN}\} $ when $ N_{COAL} $ is fixed (fig \ref{fig2:a}), and in $ \{N_{COAL}, P^{MIN}\} $ for a fixed $ \phi $ (fig \ref{fig2:b})}
\label{Fig2}
\end{figure}

As shown in figure \ref{fig2:a}, the parameters $\phi$ and $P^{MIN}$ shape the utility function such that, if $ \phi $ is close to zero, the reliability requirement is very high and only small values of $ P^{MIN}$. This could potentially lead to valid coalitions (and positive utility values). Conversely, the higher $\phi$, the less constraints are imposed to the coalitions and more valid coalitions can arise for a larger spectrum of $ P^{MIN}$. Obviously, the highest utility values are found for high $ \phi $, because the coalitions formed are able then to announce higher contract values, yielding higher utilities.

In figure \ref{fig2:b}, we fix the reliability value to a given empirical value ($\phi = 0.08 $). We can observe how the social welfare $\sum_{i \in S} \tau_i(S)$ evolves according to $\phi$ and $P^{MIN}$. In our setting we can observe  a maximum point for $ N_{COAL} = 3 $, followed a decrease. There is clearly a maximum number of coalitions sustainable for given set of parameters. After that point we will form more coalitions that would be unable to pass the grid requirements. Moreover, reckon that increasing $ N_{COAL} $ means also increasing $ \epsilon $, leading to denser graphs, meaning that the algorithm performance will also decrease.

\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.8]{./figure5/coals}
  \caption{Evolution of the coalitions for different values of $ P^{MIN} $ for clique percolation (blue dots) and random process (red dots). The diameters of the dots are proportional to the corresponding coalition sizes.}
  \label{Fig3}
\end{figure}

In figure \ref{Fig3}, we perform a simulation with 200 prosumers, and we fix the reliability $ \phi = 0.1 $ meaning that coalitions should produce more than their contract values at least $ 90 \% $ of the time. We observe the result of our algorithm for a fixed number of coalitions to $N_{coal} = 10$ and let the algorithm form the coalitions as the grid changes according to $ P^{MIN} $. We compare our algorithm with a random process that forms coalition. Figure \ref{Fig3} shows this evolution for our algorithm (blue dots) and for the random process (red dots). The diameter of a dot is proportional to the number of agents in the coalition and the higher the dot is, the higher its utility. The $ P^{MIN} $ values of the x axis are expressed in tenth of MW for readability and the hatched zone corresponds to the "under-requirement space", meaning that whenever a coalition is in this zone, it has a null utility.

Looking at figure \ref{Fig3}, we can see first that our algorithm performs better at finding high valued coalitions. The blue dots allowed to enter the market are indeed outnumbering the red dots, especially when the grid requirements are neither too low nor too high ($ P^{MIN} = 5 $ in figure \ref{Fig3}).

In more details, figure \ref{fig4:a} presents the evolution of social welfare as the number of coalitions increases (all other parameters are kept constant) for random (red curve), clique percolation (blue curve), and correlated coalitions (green curves) that stands for a worst case scenario. As for figure \ref{fig4:b}, it shows the percentage of coalitions able to enter the market for different values of $ P^{MIN} $. For consistency, we average the results of both plots of the random procedure over 100 realizations with 95\% of confidence interval.

>>>>

\begin{equation}
 \mathcal{U}_{\phi,\ P^{MIN}}(S) = \mathbf{1}_{\textit{S\ valid}} \dfrac{P_{\phi}(S)}{|S|} 
\label{utility}
\end{equation}

@@

\begin{figure}
 \centering
  \subcaptionbox{$ (\phi, P^{MIN}) $ subspace\label{fig2:a}}{\includegraphics[width=1.7in]{./figure1/util}}
  \subcaptionbox{$ (N_{COAL}, P^{MIN}) $ subspace\label{fig2:b}}{\includegraphics[width=1.70in]{./figure6/util2}}
\caption{Utility function as function of the parameters space $\{\phi \in [0,1], N_{COAL}\in \mathbb{Z}^+, P^{MIN} \in \mathbb{R}^+\}$: In $ \{\phi \in [0,0.6], P^{MIN}\} $ when $ N_{COAL} $ is fixed (fig \ref{fig2:a}), and in $ \{N_{COAL}, P^{MIN}\} $ for a fixed $ \phi $ (fig \ref{fig2:b}). For readability, $ P^{MIN} $ is expressed in tenth of MW.}
\label{Fig2}
\end{figure}

As shown in figure \ref{fig2:a}, the parameters $\phi$ and $P^{MIN}$ shape the utility function such that, if $ \phi $ is close to zero, the reliability requirement is very high. Only small values of $ P^{MIN}$ could then potentially lead to valid coalitions (and positive utility values). Conversely, the higher $\phi$, the less constraints are imposed to the coalitions and more valid coalitions can arise for a larger spectrum of $ P^{MIN}$. Obviously, the highest utility values are found for high $ \phi $, because the coalitions formed are able then to announce higher contract values, yielding higher utilities.

In figure \ref{fig2:b}, we fix the reliability value to a given empirical value ($\phi = 0.08 $). We can observe how the social welfare $\sum_{S \in CS} \mathcal{U}_{\phi,\ P^{MIN}}(S)$ evolves according to $P^{MIN}$ and $ N_{COAL} = |CS| $. In our setting we can observe  a maximum point for $ N_{COAL} = 3 $, followed by a decrease. There is clearly a maximum number of coalitions sustainable for given set of parameters. After that point we will form more coalitions that would be unable to pass the grid requirements. Moreover, reckon that increasing $ N_{COAL} $ means also increasing $ \epsilon $, leading to denser graphs, meaning that the algorithm performance will also decrease.

\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.8]{./figure5/coals}
  \caption{Evolution of the coalitions for different values of $ P^{MIN} $ for clique percolation (blue dots) and random process (red dots). The diameters of the dots are proportional to the corresponding coalition sizes. The hatched zone corresponds to the \textit{"under-requirement space"}, and results in zero utility for any coalition in this area.}
  \label{Fig3}
\end{figure}

In figure \ref{Fig3}, we perform a simulation with 200 prosumers, and we fix the reliability $ \phi = 0.1 $ meaning that coalitions should produce more than their contract values at least $ 90 \% $ of the time. We observe the result of our algorithm for a fixed number of coalitions to $N_{coal} = 10$ and let the algorithm form the coalitions as the grid changes according to $ P^{MIN} $. We compare our algorithm with a random process that forms coalition. Figure \ref{Fig3} shows this evolution for our algorithm (blue dots) and for the random process (red dots). The diameter of a dot is proportional to the number of agents in the coalition and the higher the dot is, the higher its utility. The $ P^{MIN} $ values of the x axis are expressed in tenth of MW for readability and the hatched zone corresponds to the \textit{"under-requirement space"}, meaning that whenever a coalition is in this zone, it has a null utility.

Looking at figure \ref{Fig3}, we can see first that our algorithm performs better at finding high valued coalitions. The blue dots allowed to enter the market (non hatched zone) are indeed outnumbering the red dots, especially when the grid requirements are neither too low nor too high ($ P^{MIN} = 5 $ in figure \ref{Fig3}).

In more details, figure \ref{fig4:a} presents the evolution of social welfare as the number of coalitions increases (all other parameters are kept constant) for random (red curve), clique percolation (blue curve), and correlated coalitions (green curves) that stands for a worst case scenario. As for figure \ref{fig4:b}, it shows the percentage of coalitions able to enter the market for different values of $ P^{MIN} $. For consistency, we average the results of both plots of the random procedure over 100 realizations.
